Automatic conversion of colloquial Finnish to standard Finnish

### 2015-03-28 2015-nodalida-finnormal/finnormal2015.tex
add speculation about the two rule sets
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59550 72bbbca6-d526-0410-a7d9-f06f51895060

Our evaluation corpus was created by manually translating texts in colloquial Finnish
to standard Finnish.\footnote{The corpus is freely available and published under the 
CreativeCommons {\sc cc-by-sa} 3.0 licence.} The texts were extracts from internet 
relay chat (IRC) conversations. We performed the conversion process in two steps, the first
step involved simple orthographic normalisation, for 
example \emph{oon} $\rightarrow$ \emph{olen} `I am'. Syntactic and stylistic conversions 
were not applied at this stage. The second conversion step normalised the text
both orthographically and syntactically/stylistically. Table~\ref{table:corpexample} presents
an excerpt from each of the three parts of the corpus.

@@

Out of the rule-based systems, the set of rules which was created by examining the development corpus
outperforms both the set of rules from the grammar and the combined rules.

It is interesting to note that supervised systems trained on the normalised section of the corpus outperform
those trained on the standardised corpus. One explanation for this could be that the corpus size is small, so 
that the word alignments are not as reliable on the standardised corpus which is by nature not a word-for-word 
conversion.

@@

Although a reasonable size for a test corpus, the corpus is still too small for 
wide-coverage experiments. We intend to expand the corpus size to at least 10,000
words. Another weakness of the corpus is that it contains text from a single author,
we would ideally like to add texts from other authors and other colloquial genres. Although
the challenge here is finding text that is both free of privacy issues and available
to release under a free/open-source licence.

@@

Joonas Kylm\"{a}l\"{a}

>>>>

Our evaluation corpus was created by manually translating texts in colloquial Finnish
to standard Finnish. The corpus is freely available and published under the 
CreativeCommons {\sc cc-by-sa} 3.0 licence\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-fin/texts/normalisation/}.
The texts were extracts from internet 
relay chat (IRC) conversations. We performed the conversion process in two steps, the first
step involved simple orthographic normalisation, for 
example \emph{oon} $\rightarrow$ \emph{olen} `I am'. Syntactic and stylistic conversions 
were not applied at this stage. The second conversion step normalised the text
both orthographically and syntactically/stylistically. Table~\ref{table:corpexample} presents
an excerpt from each of the three parts of the corpus.

@@

Out of the rule-based systems, the set of rules which was created by examining the development corpus
outperforms both the set of rules from the grammar and the combined rules.
The rules from the grammar capture more general tendencies, whereas the rules from the development corpus are more lexicalised.
Since the testing corpus is small and only contains text from a single author, the higher performance of the second rule set could also be an due to overfitting.

It is interesting to note that MT systems trained on the normalised section of the corpus outperform
those trained on the standardised corpus. One explanation for this could be that the corpus size is small, so 
that the word alignments are not as reliable on the standardised corpus which is by nature not a word-for-word 
conversion.

@@

Although a reasonable size for a test corpus, the corpus is still too small for 
wide-coverage experiments. We intend to expand the corpus size to at least 10,000
words. Another weakness of the corpus is that it contains text from a single author.
We would ideally like to add texts from other authors and other colloquial genres---the 
challenge here is finding text that is both free of privacy issues and available
to release under a free/open-source sence.

@@

We thank Joonas Kylm\"{a}l\"{a} for translating the colloquial sentences into normalised and standardised versions.
